|

### 2021-08-27 article/article.Rmd
article revising 1

Corruption is one of the most prevalent and harmful XXXX of developing countries. Corruption lowers public goods provision \citep{banerjee201227, beekman2014corruption}, decreases foreign investments \citep{habib2002corruption}, stunts economic growth \citep{mauro1995corruption, grundler2019corruption}, and affects human capital investments \citep{ferraz2012corrupting}. Corruption also has detrimental effects on interpersonal trust \citep{you2018trust}, attitudes toward political parties \citep{winters2015political}, and fuels the distrust in the political system \citep{mignozzetti2019brazil}. The International Transparency estimates that corruption is responsible for $1.23 trillion US dollars every year, a sum that could lift 1.4 billion people out of the extreme poverty for six consecutive years \citep{fleming2019corruption}.

@@

The game starts in the \textit{governing stage}, with the incumbent deciding the governmental investment in public goods provision. Incumbents have a budget $R$ units, and they choose the level of investment in public goods between zero and $R$. To decide about the investment, they receive privileged information from governmental agencies about the optimal investment required to maximize voters' welfare. For example, if a politician is deciding the funding for a governmental agency, the optimal budget is the amount that makes the agency to function perfectly. Less resources than needed could decrease significantly the quality of services but more resources than needed would not improve the services much. In technical terms, the optimal amount of service provision is a random variable $\theta \in [0, R]$, that is drawn before the game starts, and it is informed to the incumbent after she wins the election. I assume that the optimal level of expenditure is drawn from a uniform probability distribution, with support in $[0, R]$.\footnote{In the online companion of this paper, I solve the model for a generic probability density function $f$.}

When deciding the expenditure, the incumbent can choose to invest more or less money then the optimal investment level. If the incumbent decide to invest less money, I assume that the goods are not provided. This simplification matches the reality in most cases, especially in developing economies, where restricting investments can cripple the entirety of the service provision. On the other hand, if the incumbent proposes an expenditure higher than the required investment, I assume that she pockets the difference between the proposed and the optimal expenditure. This has two implications: first, the public goods are always provided, satisfying the voters' needs; second, the incumbent has an excuse and a method to extract rents. Although corruption that ensures public goods provision seem counter-intuitive, the whole idea of kickbacks presumes that the expenditure is materialized. For instance, during the 2016 Brazilian World Cup, the stadiums constructed with public money were delivered, despite the fact that they costed 1.5 to 2.5 times more than comparable stadiums in less corrupt countries \citep{XXXX}.

After the governing stage, it starts the \textit{campaign stage}. During the campaign, the incumbent faces a challenger in an election with two characteristics. First, I assume that voters have a incumbent-bias, so that voters prefer to reelect incumbents for sure when the public good were successfully provided. This XXXX. Second, challengers have only one campaign instrument at their disposal: to use negative campaign against the incumbent, expose their malfeasance (if present), and XXXX.

Once in office, they learn the optimal public goods provision, but they have discretion to decide   is responsible for investing in public goods provision for the voters. For instance, XXXX.

XXXX, and proposes a public goods provision XXXX. The incumbent has a fixed budget of $A$ units, to invest in service provision. The level of expenditure required to provide the public goods is a random variable $\eta \in [0, A]$, where  The randomness reflects the idea that changes in real-world conditions influence the amount needed for service provision. The incumbent knows the optimal investment level, making it her private information. If incumbents were honest, they would invest the needed amount to provide the services ($\eta$), but in this model, the incumbent invests the optimal amount plus some value, that she intends to pocket. The utility from pocketing money is described by $u(e_i-\eta)$, which $u$ is a strictly concave utility function, and $e_i$ represents the service expenditure by an incumbent $i$.\footnote{This assumption matches a frequent empirical regularity in the political environment, especially in new democracies: corruption tend to be over-investments, meaning that politicians add kickbacks and gifts to procurement contracts, receiving benefits them. Usually, what happens is not that voters fail to receive the services, it is that those services cost much more than their market value. For instance, \citet{Bandiera2009} estimated that the waste in procurement account for around 2\% of the Italian GDP.}

After the service provision period, the electoral stage starts with the incumbent and the challenger campaigning. The challenger and the voters do not observe the optimal service investment level, but the challenger observes the amount of money invested by the incumbent politician ($e_i$, for all $i$). The challenger faces no entry costs and has the opportunity to use the negative campaign to convince the voters that the incumbent has embezzled public funds.\footnote{ This assumption is strong, as many papers demonstrated the effect of quality and service provision on the decision to challenge an incumbent \citep{Gordonetal2007, ashworth2008electoral, gordon2009voter, hall2015much}. In this paper, I endogenously fix the number of incumbents and challengers, as the critical point here is to study how the number of politicians influences incentives. In the Appendix, I extend the model to have a pool of possible challengers that have their entry costs a random variable. In the extension, the total pool of likely entrants becomes the critical variable, holding all qualitative findings here unchanged. Therefore, to fix the number of entrants is easier to understand, and represents a class of models that preserve the characteristics of our results, than we adopt this approach in the main text. Also, for the entry decision in economics, see \citet{Milgrom1982, Nti1989}.} If she decides to use all of her campaign resources to expose the corrupt politician, she pays a cost $c>0$. Her chances of success depend on the strength of the corruption evidence, plus other environmental facts, such as how much control over information incumbents have \citet{ferraz2011electoral, boas2011controlling}, credibility perceived by the voter \citep{polborn2006informative, AshworthShotts2011}, access to privileged information, and the challengers' competence to spread the information among voters. I assume that these factors are captured by a corruption detection technology, a function $\lambda(.)$ that is increasing on the amount embezzled by the incumbents. It increases with the amount embezzled, as the more money the incumbent pockets, the easier to convince voters that there is hard evidence of malfeasance undertaken by the incumbents.\footnote{There are two ways to interpret the hard evidence detection technology. First, it can be exposure of hard corruption evidence that was brought by a law enforcement agency, such as the FBI in the US or the General Comptroller Agency in Brazil. Second, it can be information uncovered during the campaign process by candidates or former allies that turned to the opposition. Either way, we consider that the process is noisy, but more substantial sums of money embezzled makes its detection easier.}

Finally, as the voters' decisions are sequentially rational, upon observing hard corruption evidence, they punish the incumbent electing the challenger. Otherwise, the claim that the incumbent is corrupt is interpreted as a cheap talk by the voters.\footnote{Note that it is not ex-ante rational to always dismiss (never dismiss) the politicians, even when the voters know that politicians are corrupt. This because always dismissing (never dismissing) would generate equilibrium where politicians always spend the total amount of resources, regardless of how much is needed to provide the services.} Thus, the voter only punishes the incumbent when the challenger presents and convinces the voters that there is hard evidence of malfeasance. The game then ends, and payoffs are realized. The timeline for this model is as follows:

\begin{enumerate}
 \item Nature draws the optimal investment level to provide the public services.
 \item The incumbent observes the optimal investment level and choose the expenditure.
 \item The election starts with the challenger choosing her campaign strategy.
 \item Voters observe the campaign and choose whether to retain or dismiss the incumbent.
\end{enumerate}

The strategic decisions in this game are the expenditure level chosen by the incumbent ($\varepsilon$), the fraction of time that the challenger invests in exposing the corrupt incumbent ($\varphi$), and the voter's decision to retain or dismiss the incumbent. Condensing the discussion above, we have the following expected utility functions. For the incumbents:

\[
U_i (\varepsilon_i; \ \eta, \ \bvphi^{*}, \ \bups^{*}, \ \varepsilon_{-i}^{*}) \ = \ u(\varepsilon_i - \eta) + B \ \Prob (\upsilon_i = 1 \ | \ \psi, \dd) 
\]

Where $B$ is the reward for staying in office (assumed the same for every player) and $\Prob (\upsilon_i = 1 \ | \ \psi, \dd)$ is the probability of being retained, given $\psi$ and the exposure profile $\dd$. For a challenger $j$:

\[
U_j (\bvphi_j; \ \bvphi^{*}_{-j}, \ \bups^{*}, \ \bveps^{*}) \ = \ B \int_{0}^A \Prob (\upsilon_j = 1 \ | \ \eta, \bveps^{*}, \bvphi^{*})\mu(\eta|\bveps^{*}) d\mu(\eta) - c \sum_{i=1}^K \varphi_{ij}(\psi, \bveps^{*})
\]

With $c > 0$ being the cost of performing one full corruption exposure. Notice that the total amount spent on exposing corrupt politicians will be the sum of the amounts spent on each exposure. The representative voter's expected utility is equal to:

\[
U_V (\psi, \dd, \bups) \ = \ \indf \{\psi=G \}r_G
\]

Where $\indf$ is the indicator function for $\psi=G$ and $r_G$ the value of the public good for the voter. Notice that no $\psi=B$ is observed on the equilibrium path, as under-spending is a dominated strategy.

Any retention rule selected for the voters is sequentially rational because there is no uncertainty about the results when they cast their votes. I impose the retention rule for the voter to be to dismiss all incumbents when $\psi=B$ or replace incumbents that were successfully exposed by challengers when $\psi=G$. This rule is ex-post efficient as over-dismissal would make incumbents pocket all the money regardless of $\eta$.

I solve this game for a \textit{Perfect Bayesian Equilibrium} \citep{Fudenberg1991book, Fudenberg1991}, and as this is a signaling game, I also use \textit{universal divinity} as the equilibrium selection criteria \citep{BanksSobel1987}.\footnote{The restriction imposed by the universal divinity requires that the challengers, facing a deviation from the equilibrium path, put probability one on the type that is more likely to deviate. Moreover, as we have a continuum of types, we use the approach by \citet{Malaith1987}, which consists in finding the monotonicity in the payoffs and single crossing conditions, characterizing the behavior of the types inside the intervals of the type space, analyzing the incentives to pool or separate among the types.}

\begin{definition}
 \label{def_eqn}
 A Perfect Bayesian Equilibrium for the game described, with $|I|=K$ and $|C|=L$, $K$ and $L$ in $\mathbb{N}$, type space $T=[0,A]$, consists in optimal strategies $\bveps^{*}$, $\bvphi^{*}$, $\bups^{*}$ and beliefs $\mu^*$, that satisfy
 \begin{enumerate}
  \item $\forall t_i \in T$, $\varepsilon^{*}_i(t_i)$ maximizes $ U_i (\varepsilon_i; \ t_i, \ \bvphi^{*}, \ \bups^{*}, \ \varepsilon_{-i}^{*}) $ $\forall i \in \{1, \cdots, K\}$.
  \item $\forall \bveps \in \left[0,A \right]^K $ and $\psi \in \{B, \ G \}$, $\bvphi^{*}_j(\bveps, \psi)$ maximizes $ U_j (\bvphi_j; \ \bvphi^{*}_{-j}, \ \bups^{*}, \ \bveps^{*}, \ \mu^*) $, $\forall j \in \{1, \cdots, L\}$
  \item $\forall \psi \in \{B, \ G \}$ and $\dd \in \{0,1\}^L$, $\bups^{*}(\psi, \dd)$ maximizes $ U_V (\psi, \dd, \bups^*) $.
  \item The beliefs are derived using Bayes rule whenever possible, and for all $\varepsilon_i$ that are off-the-equilibrium path, $\mu^*(t_i| \varepsilon_i)>0$ (in fact, 1) only if $\widehat{\varepsilon}_i \in \arg \min_{t_i} p(t_i, \widehat{\varepsilon}_i)$, where $ p(t_i, \widehat{\varepsilon}_i)$ is the profit from deviating from $\varepsilon_i$ to $\tilde{\varepsilon}_i$ for a type $t_i$.
 \end{enumerate}
\end{definition}

Note four characteristics of the model proposed. First, the incumbent politicians always embezzle some amount in equilibrium. Second, conditional on the benefits from holding office, to spend less than the optimal investment is harmful to the incumbent. In concrete terms, this means that politicians' first aim is to be elected. If the voter always dismisses the politician, then the politician will always say that the good costs $A$. Third, the voters always receive the public goods, regardless of the amount of money embezzled. The voter decides to punish the corrupt incumbent to discipline the incumbent's behavior, minimizing the rents extracted. Finally, this model has no ideological proximity, or affinity between the voter and any of the politicians: the paper aims to explore the incentives for corruption, not the trade-offs between ideology and corruption. Ideology is vital to understand why voters may not want to punish a corrupt politician, but this is not the main point of this paper, and adding this to the paper only blurs the logic of the argument. I now solve the model for the scenarios relevant to this paper.

\subsection{One incumbent and one challenger}

>>>>

Corruption is one of the most prevalent and harmful XXXX of developing countries. 2B2

@@

The game starts in the \textit{governing stage}, with the incumbent deciding the governmental investment in public goods provision. Incumbents have a budget $R$ units, and they choose the level of investment in public goods between zero and $R$. To decide about the investment, they receive privileged information from governmental agencies about the optimal investment required to maximize voters' welfare. For example, if a politician is deciding the funding for a governmental agency, the optimal budget is the amount that makes the agency to function perfectly. Less resources than needed could decrease significantly the quality of services but more resources than needed would not improve the services much. In technical terms, the optimal amount of service provision is a random variable $\theta \in [0, R]$, that is drawn before the game starts, and it is informed to the incumbent after she wins the election. I assume that the optimal level of expenditure is drawn from a uniform probability distribution, with support in $[0, R]$.\footnote{In the online companion of this paper, I solve the model for a generic probability density function $f$, and lay down the conditions for an equilibrium.}

When deciding the public goods provision, the incumbent can choose to invest more or less money then the optimal investment level. If the incumbent decides to invest less money, I assume that the goods are not provided. This simplification matches the reality in most cases where restricting investments can cripple the entirety of the service provision. Conversely, when the incumbent proposes an expenditure higher than the required investment, I assume she pockets the difference between the proposed and the optimal expenditure. This assumption has two implications: first, the public goods are always provided, satisfying the voters' needs; second, the incumbent has an excuse and a method to extract rents. Although corruption that ensures public goods provision seem counter-intuitive, the whole principle of kickbacks presumes that the service provision is materialized. For instance, during the 2016 Brazilian World Cup, the stadiums constructed with public money were delivered for the world cup, despite the fact that they costed 1.5 to 2.5 times more than comparable stadiums in less corrupt countries \citep{XXXX}.

After the governing stage, it starts the \textit{campaign stage}. During the campaign, the incumbent faces a challenger in an election with two characteristics. First, voters have a incumbent-bias, so that they prefer to reelect incumbents for sure when the public goods were successfully provided. This correlates with incumbency advantage findings \citep{XXXX} as well as formal models of public service provision \citep{XXXX}. Second, I assume that challengers have only one campaign instrument at their disposal: to use negative campaign against the incumbent, expose their malfeasance if present, and receive the benefits from the denouncing. Although this seems unrealistic, having only one campaign strategy helps us focus on to what extent this strategy will be employed. Moreover, when the challenger decides to run negative campaign against an incumbent, she pays a value $c>0$. This value captures the direct costs of collecting information about corruption, and also XXXX.

Finally, the last stage of the game features the representative voter observing the information gathered by the challengers, and deciding whether to sanction the incumbent. Challengers have an incentive to run negative campaign against incumbents, regardless of whether the incumbent engaged in corruption. The voters, knowing the challenger's incentives to falsely accuse incumbents, evaluate the presented evidence. The strength of the evidence correlates with the difference between the optimal expenditure and the proposed expenditure. Voters and challengers do not know the optimal expenditure, but the greater the distance between the optimal and the enacted expenditures, the easier for challengers to convince voters that the incumbents are corrupt. This assumption has empirical support is most accounting literature, as the higher the kickbacks, the harder it is to hide it from authorities \citep{XXXX}.

The timeline for this model is as follows:

\begin{enumerate}
 \item Nature draws the optimal investment level to provide the public services.
 \item The incumbent observes the optimal investment level and chooses the government investment in public goods.
 \item After the public goods provision, the election starts with the challenger choosing her campaign strategy.
 \item Voters observe the campaign and choose whether to retain or dismiss the incumbent.
\end{enumerate}

To differentiate between strategies and actions, I assume that the incumbent's decision for the expenditure is a function $\varepsilon: [0,R] \rightarrow [0,R]$, i. e., the decision-making process receives the information about the optimal level of public goods provision, and returns the expenditure decision. For the challenger there is a similar reasoning, and the decision to investigate the incumbent is a function $\varphi: [0,R] \rightarrow [0,1]$, that receives the expenditure decision and returns a probability of running the negative campaign. Condensing the discussion above, the incumbent, upon observing the optimal public goods provision level $\theta \in [0,R]$, decides the expenditure $\varepsilon \in [0,R]$. The benefits from pocketing resources are captured by a function $u(.)$. In the main paper, I assume it is equal to the identity function but that in the appendix I solve the model for a generic concave function. The expected utility of the incumbent also adds the benefit from winning the election $B>0$ times the probability of this event, given the optimal and enacted expenditures and the decision to investigate $\varphi$.

\[
U_I (\varepsilon; \theta, \varphi) \ = \ u(\varepsilon - \theta) + B \mathbb{P} (\text{re-election}; \varepsilon, \theta, \varphi) 
\]

The expected utility for the challenger involves a decision in terms of the probability to win the election and the costs of running a negative campaign. This benefit from holding office  $B>0$ is multiplied by the probability that the challenger replaces the incumbent.\footnote{I assume the material and non-material benefits from office the same for both the challenger and the incumbent. This simplifies the exposition, and makes it easier to think about these benefits as salaries or other types of pecuniary benefits.} The probability that the challenger wins the election depend on the decision to run a negative campaign, the enacted expenditure, and on her beliefs about the optimal expenditure, when she observes the enacted expenditure. The beliefs are captured by a probability rule $\mu(\theta|\varepsilon)$. The challenger also decides the investigation strategy $\varphi \in [0,R]$, and if she undertakes the strategy, then she pays a cost of $c>0$.

\[
U_C (\varphi; \varepsilon) \ = \ B \int_{0}^R \mathbb{P}(election|\varphi, \varepsilon, \theta) \mu(\theta|\varepsilon) d\mu(\theta) - c \varphi
\]

Finally, the voter's decisions are sequentially rational, in the sense that they are undertaken after the expenditure and investigations are undertaken. They cannot change the decisions ex-post, and this facilitates the decision rules for voters. Ex-ante, I assume that the best is to remove from office the incumbent whenever the challenger credibly shows evidence that the incumbent is corrupt. The chance that the challenger convinces the voter is a function $d(\varepsilon-\theta)$, and to simplify, I consider that this function is equal to $d(\varepsilon-\theta) = \dfrac{\varepsilon-\theta}{R}$. In the appendix, I model the decision for a generic functional form for $d(.)$.

I solve this game for a \textit{Perfect Bayesian Equilibrium} \citep{Fudenberg1991book, Fudenberg1991}, and as this is a signaling game, I also use \textit{universal divinity} as the equilibrium selection criteria \citep{BanksSobel1987}. The restriction imposed by the universal divinity requires that the challenger, facing a deviation from the equilibrium path, assign a probability one on for the type that is more likely to deviate. Moreover, as I have a continuum of possible investment strategies, I use the approach by \citet{Malaith1987}, which consists in finding the monotonicity in the payoffs and single crossing conditions, characterizing the behavior of the types inside the intervals of the type space, analyzing the incentives to pool or separate among the types.

\begin{definition}
 \label{def_eqn}
 A Perfect Bayesian Equilibrium for the strategic interaction described consists in strategic choices of the expenditure $\varepsilon$ and the negative campaign $\varphi$ such that:
 \begin{enumerate}
  \item For an optimal expenditure level, $\theta \in [0,R]$, the proposed expenditure $\varepsilon$ maximizes $U_I (\varepsilon; \varphi, \theta)$.
  \item $\forall \bveps \in \left[0,A \right]^K $ and $\psi \in \{B, \ G \}$, $\bvphi^{*}_j(\bveps, \psi)$ maximizes $ U_j (\bvphi_j; \ \bvphi^{*}_{-j}, \ \bups^{*}, \ \bveps^{*}, \ \mu^*) $, $\forall j \in \{1, \cdots, L\}$
  \item The beliefs are derived using Bayes Rule, whenever possible. The universal divinity also requires that for all expenditure levels that are off-the-equilibrium path, $\mu(\theta| \varepsilon)>0$ (in fact, 1) only if $\widehat{\varepsilon} \in \arg \min_{\theta} p(\theta, \widehat{\varepsilon})$, where $ p(\theta, \widehat{\varepsilon})$ is the profit from deviating from the optimal $\varepsilon$ to a deviation expenditure $\tilde{\varepsilon}$, when the optimal expenditure level is $\theta$.
 \end{enumerate}
\end{definition}

Note that this model has no ideological proximity, or affinity between the voter and any of the politician. This because the paper's objective is to explore the incentives for corruption, not the trade-offs between ideology and corruption.

\subsection{The Solution for the One Incumbent - One Challenger Model}
